#**Reproducible Research: Peer Assessment 1**

```{r,results='hide',message=FALSE,warning=FALSE}
library(Hmisc)
library(ggplot2)
```


## Loading and preprocessing the data


```{r}
unzip('activity.zip')
data<-read.csv("activity.csv")
```


## What is mean total number of steps taken per day?

```{r} 
stepsbyDay<- aggregate(steps ~ date, FUN = sum, data)

hist(stepsbyDay$steps,col = "red",main = "Total number of steps each day during October and November 2012",xlab = "Total steps")

Mean<-mean(stepsbyDay$steps)
Median<-median(stepsbyDay$steps)
```

Mean is **`r Mean`** and the median is **`r Median`**.
## What is the average daily activity pattern?

```{r}
stepsbyInterval <- aggregate(steps ~ interval, data, mean)

plot(stepsbyInterval$interval,stepsbyInterval$steps, type="l", xlab="5-minute interval", ylab="no. of steps averaged  across all days",main="Average steps taken everyday for each 5-minute interval")

Max_steps_interval <- stepsbyInterval[which.max(stepsbyInterval$steps),1]
```

The interval with maximum number of steps is **`r Max_steps_interval`**.

## Imputing missing values

```{r}
no_of_missing<-sum(!complete.cases(data))

```

No of missing values is **`r no_of_missing`**

```{r}
new_data<-data
new_data$steps<-impute(new_data$steps,mean)
stepsbyDayi<- aggregate(steps ~ date, FUN = sum, new_data)

hist(stepsbyDayi$steps,xlab="Total no. of steps taken each day",col="blue")

Mean_imputed<-mean(stepsbyDayi$steps)
Median_imputed<-median(stepsbyDayi$steps)
diff1<-Mean_imputed-Mean
diff2<-Median_imputed-Median
steps_diff<-sum(stepsbyDayi$steps)-sum(stepsbyDay$steps)
```
The imputed mean and median are **`r Mean_imputed`** and **`r Median_imputed`**  
Mean and median differences are **`r diff1`** and **`r diff2`**  
Total difference in steps:**`r steps_diff`**

## Are there differences in activity patterns between weekdays and weekends?

``` {r}
new_data$date <- as.Date(new_data$date)
weekdays <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
new_data$wDay <- factor((weekdays(new_data$date) %in% weekdays),levels=c(FALSE, TRUE), labels=c('weekend', 'weekday'))

stepsbyInterval<-aggregate(steps ~ interval + wDay, new_data, mean)

ggplot(stepsbyInterval,aes(interval,steps))+geom_line(col="blue")+xlab("Interval")+ylab("NUmber of steps")+facet_grid(wDay ~ .)


``` 
